Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study

### 2025-05-04 docs/response/response.qmd

---
title: "Author Response to Reviewer Comments"
# format:
#   pdf: default
#bibliography: references.bib
engine: knitr
---

1.  The choice of W (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?

<!-- 2.  While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist? -->

<!-- 3.  The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model). Have the authors examined how sensitive the model results are to changes in $\rho$? -->

<!-- 4.  The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants? Would spatial effect maps or additional diagnostic tools enhance this understanding? -->

<!-- 5.  Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2). Before reaching this conclusion, did you compare GWR to other spatial regression models? -->

<!-- 6.  The scope of properties analysed includes residential houses and commercial real estate (page 16). How do the authors reconcile this with the abstract, which mentions only house price data? -->

<!-- 7.  After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16). Why does the distribution of these data not appear on the map (Figure 6)? -->

<!-- 8.  The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17). How was this approach implemented, given that there is no explanation of how each observation was georeferenced? -->

<!-- 9.  Why is Table 5 not referred to in the text? -->

<!-- 10. The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18). However, for the Lombok House Price Data, the authors have used point-level models. How do you reconcile this contradiction? -->

<!-- 11. In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results. Would the findings remain consistent if a different spatial weights structure was used? -->

<!-- 12. The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)? Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions? -->

### 2025-05-04 response/response.qmd

---
title: "Response to Journal's 2nd Reviewer Comments"
subtitle: "Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study"
author: 
  - name: Indira Puteri Kinasih
    orcid: 0009-0009-6959-6723
    email: indiraputeri@uinmataram.ac.id
---

1.  **The choice of $W$ (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?**

    Thank you for this insightful comment regarding the specification of the spatial weight matrix, $W$. We fully acknowledge that the choice of $W$—whether binary adjacency, row-standardized, or distance-based—can substantially influence the performance and interpretation of spatial models. In the present study, our primary objective was to evaluate and compare the performance of CAR, SAR, and GWR models using both simulated data and real-world data from Lombok. To maintain consistency across models and to isolate the structural differences in model behavior, we have, at this stage, limited our analysis to a binary adjacency matrix for $W$. We agree that exploring alternative specifications of $W$ (e.g., row-standardized or distance-decay matrices) is a valuable direction for further research. We plan to extend our analysis in future work to assess how varying spatial weight structures affect the robustness and predictive accuracy of these spatial models.

2.  **While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist?**

    Thank you for your thoughtful question. We acknowledge that in small-area data, bias can indeed become a concern due to limited information at the unit level. However, in the case of the multilevel CAR (mlvCAR) model, each area contains multiple observations, allowing us to model area-specific effects—functionally resembling a random effects structure. This hierarchical structure helps mitigate the impact of sparsity within individual areas.

    Regarding sample size, prior studies suggest that having approximately 15–30 observations per area is generally sufficient for stable estimation in multilevel frameworks. Given our data structure aligns with this recommendation, we believe the mlvCAR model offers a robust approach, especially in real-world scenarios where heterogeneity and partial pooling across areas are important. We thus recommend the use of mlvCAR to address both the potential bias and boundary effects effectively.

3.  **The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model). Have the authors examined how sensitive the model results are to changes in $\rho$?**

    In our study, we recognize that the value of $\rho$ plays a crucial role in determining the extent of spatial autocorrelation captured by the model. While we did not conduct a dedicated sensitivity analysis across a range of fixed $\rho$ values, our simulation framework inherently explores variation in $\rho$ through repeated data generation processes, in which $\rho$ is estimated as part of the model fitting. The observed variability in $\rho$ reflects the underlying spatial structure in the data rather than instability in the model.

    That said, we agree that a more systematic sensitivity analysis—such as fitting the model under different fixed values of $\rho$—could offer further insights into model robustness. We consider this a valuable direction for future work, especially in scenarios with varying levels of spatial autocorrelation or data sparsity.

4.  **The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants? Would spatial effect maps or additional diagnostic tools enhance this understanding?**

    Thank you for highlighting the importance of interpreting the spatial effects $\phi$ in the CAR model. We agree that while $\phi$ captures unobserved spatial heterogeneity, its interpretability in the context of house price determinants requires careful consideration. Rather than directly reflecting specific covariates, $\phi$ accounts for spatially structured residual variation—such as omitted neighborhood characteristics, local amenities, or spatial spillovers—that are not captured by observed variables.

    To aid interpretation, we acknowledge that visualizing $\phi$ through spatial effect maps could be a powerful diagnostic tool. Such maps can reveal clusters of under- or over-prediction and provide insights into areas where additional covariates or local effects might be relevant. In the revised version of this study, we will include a Lombok sub-district-level map displaying the estimated $\phi$ values to better communicate the spatial structure captured by the model and to guide further interpretation and refinement.

5.  **Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2). Before reaching this conclusion, did you compare GWR to other spatial regression models?**

    Thank you for this thoughtful question. In our manuscript, we highlight Geographically Weighted Regression (GWR) as a prominent method in the literature, particularly due to its frequent application in house price studies. This observation is based on a review of previous works rather than a performance-based comparison. While GWR is indeed widely used for its ability to model spatial heterogeneity at the local level, our study goes further by systematically comparing GWR with CAR and SAR models through simulation and real data applications. This comparative analysis aims to evaluate their relative strengths and limitations more objectively.

6.  **The scope of properties analysed includes residential houses and commercial real estate (page 16). How do the authors reconcile this with the abstract, which mentions only house price data?**

     We acknowledge that the description in the manuscript may not have been sufficiently clear. The original dataset obtained through web-scraping did include a broad range of property types—such as residential houses, villas, boarding houses, and vacant land. However, for the purposes of this study, we restricted our analysis to residential houses only. We will revise the manuscript accordingly to ensure that this scope is clearly stated in both the abstract and the data collection section.

7.  **After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16). Why does the distribution of these data not appear on the map (Figure 6)?**

    For clarity, we note that the Lombok house price dataset used in this study does not contain precise geographic coordinates for each observation. As a result, it is not possible to accurately display the distribution of the data points on a map. While each observation is linked to a sub-district, the absence of point-level location data limits our ability to visualize them spatially within those areas. Therefore, for a more informative visualization, we are preparing a heatmap of house prices at the sub-district level, which will reflect the spatial variation in prices across Lombok using aggregated data.

8.  **The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17). How was this approach implemented, given that there is no explanation of how each observation was georeferenced?**

    Thank you for this important observation. We clarify that GLMM and multilevel CAR models do not require explicit geographic coordinates for each individual observation. Instead, they rely on an area-level identifier (ID) that links each observation to a corresponding spatial unit (e.g., sub-district).

    In contrast, GWR requires spatial coordinates for each observation. To meet this requirement, we generated random points around the centroid of each sub-district, using `st_sample()`, where the number of points corresponds to the number of observations in that area. Each point was placed within a 0.5 km radius from the centroid, ensuring a reasonable spatial spread while maintaining the area-specific context. This approach enabled us to simulate realistic spatial positions for property transactions in the absence of precise geolocation data.

9.  **Why is Table 5 not referred to in the text?**

    In the revised version of the manuscript, we have made sure to explicitly refer to Table 5 in the main text to improve clarity and consistency. We appreciate your careful reading and helpful suggestion.

10. **The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18). However, for the Lombok House Price Data, the authors have used point-level models. How do you reconcile this contradiction?**

    As we noted in our response to Comment 8, the Lombok house price data do not come with precise geographic coordinates. In order to apply the `gwr()` function, which requires point-level location data, we generated synthetic coordinates by simulating random points around the centroid of each sub-district. While this workaround allowed us to run the GWR model, we acknowledge that it is not an ideal or practical solution for real-world applications. This limitation reinforces our argument that point-level models such as GWR can be challenging to apply meaningfully when precise location data are not available.

11. **In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results. Would the findings remain consistent if a different spatial weights structure was used?**

    As mentioned in our response to Comment 1, in this study we used a binary adjacency matrix to define spatial relationships, primarily to maintain consistency across the CAR and SAR models and to focus on comparing their general performance. We acknowledge that the choice of spatial weights matrix ($W$) can significantly influence model outcomes. While we have not conducted a sensitivity analysis in this current version, we agree that testing alternative specifications—such as row-standardized or distance-based matrices—would be valuable. We consider this an important direction for future work to assess the robustness of our findings under different definitions of spatial connectivity.

12. **The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)? Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions?**

    Thank you for this important point. Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater. In contrast, there is relatively limited research focused on low-density or semi-rural contexts such as Lombok. Our study aims to address this gap by evaluating the applicability and performance of spatial models, particularly the CAR model, in a setting that has different data characteristics and spatial structures compared to major metropolitan areas.

    We also acknowledge that in richer datasets—common in urban contexts where more granular, high-frequency, and multidimensional data are available—machine learning approaches may offer competitive or even superior predictive performance. However, our focus here is to assess interpretable spatial models in contexts where data limitations remain a challenge.

### 2025-05-05 docs/_tex/manuscript.tex
update

% author setup
\author[1,2]{\fnm{Indira Puteri} \sur{Kinasih}}\email{indiraputeri@uinmataram.ac.id}\author*[1]{\fnm{Haziq} \sur{Jamil}}\email{haziq.jamil@ubd.edu.bn}\author*[1]{\fnm{Elvynna} \sur{Leong}}\email{elvynna.leong@ubd.edu.bn}
% affil setup
\affil[1]{\orgdiv{Mathematical Sciences, Faculty of
Science}, \orgname{Universiti Brunei
Darussalam}, \orgaddress{\street{Universiti Brunei Darussalam, Jalan
Tungku Link}, \city{Bandar Seri Begawan}, \postcode{BE
1410}, \country{Brunei}}}
\affil[2]{\orgdiv{Mathematics Education, Faculty of
Education}, \orgname{Universitas Islam Negeri Mataram}}

@@

Another specific ingredient in spatial modelling is the existence of
\(W\) matrix, also known as the spatial weights matrix. It encodes the
spatial relationships between areas in a study region. The weight matrix
visually shows how different regions are connected to each other,
indicating their neighbourhood relationships (Morris et al. (2019)). To
demonstrate this concept, let's take a simple example of a map with four
regions \((R_1, R_2, R_3, R_4)\), as visualised in
Figure~\ref{fig-neighsimple}.

@@

\includegraphics[width=0.7\textwidth,height=\textheight]{figures/maprec.png}

@@

This straightforward neighbourhood structure creates a single connected
component with the following relationships:
\((R_1 \sim R_2,\ R_2 \sim R_3,\ R_3 \sim R_4)\). Therefore the weight
matrix \(\boldsymbol{W}\) is

\[
\begin{array}{ccccc}
    & R_1 & R_2 & R_3 & R_4 \\
R_1 & 0 & 1 & 0 & 0\\ 
R_2 & 1 & 0 & 1 & 0\\ 
R_3 & 0 & 1 & 0 & 1\\
R_4 & 0 & 0 & 1 & 0\\
\end{array}
\]

Later on, in the modelling stage, we will also need to construct a
diagonal matrix \(\boldsymbol{D}\). It is an \(N \times N\) matrix where
each diagonal element \(d_{ii}\) denotes the number of neighbours of
region \(R_i\), while all non-diagonal elements are zero. This matrix is
essential in spatial econometric models because it captures the local
neighbourhood structure of each region. In the context of Geographically
Weighted Regression (GWR), Spatial Autoregressive (SAR), and Conditional
Autoregressive (CAR) models, \(\boldsymbol{D}\) plays a pivotal role in
defining spatial relationships and dependencies among different regions.

@@

\[
\begin{array}{ccccc}
   & R_1 & R_2 & R_3 & R_4 \\
R_1 & 1 & 0 & 0 & 0\\ 
R_2 & 0 & 2 & 0 & 0\\ 
R_3 & 0 & 0 & 2 & 0\\
R_4 & 0 & 0 & 0 & 1\\
\end{array}
\]

This matrix format indicates that region \(R_1\) has one neighbour,
regions \(R_2\) and \(R_3\) each have two neighbours, and region \(R_4\)
has one neighbour. By integrating this matrix into our modelling
framework, we ensure that the spatial structure and interactions among
regions are accurately represented, thereby enhancing the robustness and
precision of our spatial analysis.

@@

The bias values show that the models exhibit varying levels of accuracy
across the different parameters and datasets. The coefficients
\(\beta_k\) for \(k = 0, 1, 2, 3, 4\) represent the values associated
with the intercept, land size, building size, number of bedrooms, and
number of bathrooms, respectively. For \(\beta_0\), the GLM and CAR
models display relatively small biases with narrow standard deviations,
indicating better consistency, while the SAR model demonstrates larger
bias values, suggesting potential issues with accurately estimating the
intercept. For higher parameters like \(\beta_2\) and \(\beta_3\), SAR
generally has larger bias and variability, especially in the
\(SAR + CAR\) dataset, suggesting that the combined spatial structure
poses challenges for SAR's estimation. It is important to note that the
Figure~\ref{fig-biasarealevel} utilises area-level data. In this
context, the CAR model consistently demonstrates the smallest bias
compared to the other two models.

\phantomsection\label{cell-fig-biasarealevel}
\begin{figure}[H]

@@

\includegraphics[width=0.75\textwidth,height=\textheight]{figures/bias1b_plot.pdf}

@@

\caption{\label{fig-biasarealevel}Bias values by method and dataset type
for area-level data. This figure compares the bias estimates of
different methods (CAR, GLM, SAR) across datasets generated under CAR,
GWR, and SAR+CAR frameworks. Each method's performance is visualised for
four coefficients (\(\beta_1\), \(\beta_2\), \(\beta_3\), \(\beta_4\)),
highlighting variations in bias and the effectiveness of each model in
addressing spatial dependencies.}

@@

\phantomsection\label{cell-fig-biaspointlevel}
\begin{figure}[H]

\includegraphics[width=0.75\textwidth,height=\textheight]{figures/bias2_plot.pdf}

\caption{\label{fig-biaspointlevel}Bias values by method and dataset
type for point-level data. This figure compares the bias estimates of
different methods (GLMM, GWR, and mlvCAR) across datasets generated
under CAR, GWR, and SAR+CAR frameworks. Each method's performance is
visualised for four coefficients (\(\beta_1\), \(\beta_2\), \(\beta_3\),
\(\beta_4\)), highlighting variations in bias and the effectiveness of
each model in addressing spatial dependencies.}

As shown in the second plot of Figure~\ref{fig-biaspointlevel}, this
analysis utilises point-level data. Overall, the mlvCAR model
demonstrates a consistently lower bias across most parameters and
scenarios. In the \(SAR + CAR\) scenario, where spatial complexity is
most pronounced, mlvCAR achieves notably low bias, particularly for
parameters \(\beta_1\) through \(\beta_4\), indicating its suitability
for handling mixed spatial structures. making it a solid choice for a
range of spatial modelling tasks.

The GWR model, which provides highly localised estimates, displays
slightly higher bias for several parameters compared to mlvCAR,
especially for \(\beta_1\), \(\beta_2\), and \(\beta_3\) within the
\(CAR\) scenario. This suggests GWR's greater sensitivity to spatial
patterns without incorporating neighbourhood-based effects. In contrast,
GLMM, which lacks explicit spatial components, shows larger bias
variability, particularly in scenarios with pronounced spatial
dependencies, highlighting its limitations for purely spatial analyses.

@@

Table~\ref{tbl-rmse} provide insights into model accuracy at different
spatial levels. For area-level data, the models GLM, SAR, and CAR are
compared, with the CAR model consistently achieving the lowest RMSE.
This suggests that CAR is more effective at capturing spatial
dependencies at the area level. For point-level data, the models GLMM,
GWR, and mlvCAR are evaluated, with mlvCAR showing the lowest RMSE
across all datasets. This indicates that mlvCAR is particularly
well-suited for handling detailed spatial variations at the point level.
Overall, these results highlight the strengths of CAR and mlvCAR in
reducing error within their respective spatial contexts, underscoring
their potential utility in spatial modeling applications.

@@

In addition to the point-level dataset (\texttt{ppd}), which contains
individual property listings, the imputed dataset was also aggregated to
the sub-district level by calculating the mean values of key variables.
This aggregation produced an area-level dataset (\texttt{ppd\_kc}),
allowing the analysis to be conducted at both the individual and
administrative levels. These two levels of data granularity provide
complementary perspectives for evaluating the performance of spatial
models in capturing local property market dynamics.

@@

Standardization of the predictors ensures that the resulting
coefficients are on a comparable scale, facilitating meaningful
interpretation and comparison across different modeling approaches. The
results presented in the following tables provide a comprehensive
evaluation of each model's performance in capturing both the structural
and spatial characteristics of the property market in Lombok. This
comparison allows us to assess the extent to which spatially explicit
models (e.g., GWR, CAR, SAR) offer improvements over conventional
methods in explaining price variability across regions. --- Let me know
if you'd like to include metrics such as R², AIC, or Moran's I in the
paragraph as well.

We conducted model fitting on the area-level data, with detailed results
provided in the Appendix (Table~\ref{tbl-modelcomp2}). However, this
method does not provide optimal parameter estimates because data
aggregation can reduce variability and obscure spatial details essential
for capturing local effects. To address this, we used point-level data
to apply three different models: GLMM, GWR, and multilevel CAR.

@@

@@

The findings from simulation consistently indicate that both the CAR and
mlvCAR model outperform the other models in capturing spatial
dependencies. This is evidenced by lower RMSE values and reduced bias
across various dataset scenarios, highlighting the CAR model's
robustness. In datasets with strong spatial structures, such as those
generated under \(CAR\) and \(SAR + CAR\) conditions, the CAR model
achieves substantially lower RMSE values compared to GLM and SAR,
suggesting its effectiveness in handling localized spatial variations.

@@

@@

@@

::: \{.appendices latex=true\}

\section{Simple Feature Data Slice}\label{secA1}

\begin{longtable}[t]{rl}

\caption{\label{tbl-geoartfex}a slice of a simple feature collection
with 6 polygon features and 1 attribute field, area(sqm), representing
polygon sizes in square units. The dataset, using the WGS 84 CRS, spans
longitudes 116.0355--116.5 and latitudes -8.500079---8.000057. The
geometry field defines each polygon's spatial boundaries, representing
artificial study regions for spatial modeling in this research}

\toprule
area & geometry\\
\midrule
17616519 & POLYGON ((116.2296 -8.50007...\\
11485356 & POLYGON ((116.0778 -8.32479...\\
4755629 & POLYGON ((116.399 -8.380894...\\
15959450 & POLYGON ((116.3567 -8.00006...\\
12727929 & POLYGON ((116.14 -8.292773,...\\
\addlinespace
8848979 & POLYGON ((116.5 -8.210459, ...\\
7248031 & POLYGON ((116.4308 -8.45091...\\
16906811 & POLYGON ((116.1352 -8.26696...\\
6346890 & POLYGON ((116.0723 -8.30335...\\
7620652 & POLYGON ((116.4541 -8.43327...\\
\addlinespace
10987363 & POLYGON ((116.4982 -8.06795...\\
10411472 & POLYGON ((116.1071 -8.07289...\\
20452118 & POLYGON ((116.4233 -8.12744...\\
19613075 & POLYGON ((116 -8.5, 116 -8....\\
9252319 & POLYGON ((116.4541 -8.43327...\\
\bottomrule

\section{Bias values for Area-level Data}\label{secA2}

@@

[!h]
\centering\begingroup\fontsize{9}{11}\selectfont

@@

\section{Model Comparison for area-level data}\label{secA3}

@@

\caption{\label{tbl-modelcomp2}This table compares the performance of
GLM, SAR, and CAR models on area-level data. It reports parameter
estimates (with 95\% confidence intervals), spatial parameters
(\(\rho\)), variances (\(\nu^2\), \(\tau^2\)), and model fit criteria
(AIC, DIC, WAIC, LMPL, log-likelihood). The results highlight each
model's capability to capture spatial dependencies and variability}

@@

\section{Example of beta-values in GWR model}\label{secA4}

>>>>

% author setup
\author*[1,2]{\fnm{Indira Puteri} \sur{Kinasih}}\email{indiraputeri@uinmataram.ac.id}\author[1]{\fnm{Haziq} \sur{Jamil}}\email{haziq.jamil@ubd.edu.bn}\author[1]{\fnm{Elvynna} \sur{Leong}}\email{elvynna.leong@ubd.edu.bn}
% affil setup
\affil[1]{\orgdiv{Mathematical Sciences, Faculty of
Science}, \orgname{Universiti Brunei
Darussalam}, \orgaddress{\street{Universiti Brunei Darussalam, Jalan
Tungku Link}, \city{Bandar Seri Begawan}, \postcode{BE
1410}, \country{Brunei}}}
\affil[2]{\orgdiv{Mathematics Education, Faculty of
Education}, \orgname{Universitas Islam Negeri Mataram}}

@@

Another specific ingredient in spatial modelling is the existence of
\(W\) matrix, also known as the spatial weights matrix. It encodes the
spatial relationships between areas in a study region. The weight matrix
visually shows how different regions are connected to each other,
indicating their neighbourhood relationships (Morris et al. (2019)). To
demonstrate this concept, let's take a simple example of a map with 5
regions \((A, B, C, D, E)\), as visualised in
Figure~\ref{fig-neighsimple}.

@@

\includegraphics[width=0.5\textwidth,height=\textheight]{figures/sartf_plot.png}

@@

From the Figure~\ref{fig-neighsimple} we can derive an adjacency matrix
which represents the neighborhood structure of five spatial units
labeled \(A\) to \(E\). Each entry 1 indicates a direct adjacency (areas
share a common border), while 0 denotes no direct connection. For
example, Area \(A\) is adjacent to all other areas (\(B\), \(C\), \(D\),
and \(E\)), whereas Area \(B\) is only adjacent to \(A\) and \(C\). This
neighbourhood configuration captures a clear spatial interaction pattern
among the areas, which forms the basis for constructing a spatial weight
matrix, commonly denoted as the \(\mathbf{W}\) matrix

\[
\begin{array}{cccccc}
    & A & B & C & D & E \\
A & 0 & 1 & 1 & 1 & 1\\ 
B & 1 & 0 & 1 & 0 & 0\\ 
C & 1 & 1 & 0 & 1 & 0\\
D & 1 & 0 & 1 & 0 & 1\\
E & 1 & 0 & 0 & 1 & 0\\
\end{array}
\]

Later on, in the modelling stage, we will also need to construct a
diagonal matrix \(\boldsymbol{D}\). It is an \(N \times N\) matrix where
each diagonal element \(d_{ii}\) denotes the number of neighbours of
region. while all non-diagonal elements are zero. This matrix is
essential in spatial econometric models because it captures the local
neighbourhood structure of each region. In the context of Geographically
Weighted Regression (GWR), Spatial Autoregressive (SAR), and Conditional
Autoregressive (CAR) models, \(\boldsymbol{D}\) plays a pivotal role in
defining spatial relationships and dependencies among different regions.

@@

\[
\begin{array}{cccccc}
   & A & B & C & D & E \\
A & 4 & 0 & 0 & 0 & 0\\ 
B & 0 & 2 & 0 & 0 & 0\\ 
C & 0 & 0 & 3 & 0 & 0\\
D & 0 & 0 & 0 & 3 & 0\\
E & 0 & 0 & 0 & 0 & 2\\
\end{array}
\]

This matrix format indicates that region \(A\) has 4 neighbour, regions
\(C\) and \(D\) each have 3 neighbours, and region \(B\) and \(E\) has 2
neighbour.

@@

The bias values are summarised in Table~\ref{tbl-bias1} and
Table~\ref{tbl-bias2}. It shows that the models exhibit varying levels
of accuracy across the different parameters and datasets.
Figure~\ref{fig-biasplot} presents the bias estimates of regression
coefficients for different modelling approaches, separated by data type:
(a) area-level data and (b) point-level data. Each panel corresponds to
a data-generating mechanism (CAR, GWR, or SAR+CAR), and bias values are
shown with 95\% confidence intervals.

\begin{figure}

\begin{minipage}{0.50\linewidth}

@@

\includegraphics{figures/bias1_plot.png}

@@

\subcaption{\label{fig-biasplot1}Bias values by method and dataset type
for area-level data}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\includegraphics{figures/bias2_plot.png}

\subcaption{\label{fig-biasplot2}Bias values by method and dataset type
for point-level data}

\end{minipage}%

\caption{\label{fig-biasplot}Bias estimates comparison of six different
methods across datasets generated under CAR, GWR, and SAR+CAR
frameworks. Each method's performance is visualised for \(\beta_1\),
\(\beta_2\), \(\beta_3\), and \(\beta_4\), highlighting variations in
bias and the effectiveness of each model in addressing spatial
dependencies.}

@@

For area-level data (Figure~\ref{fig-biasplot1}), CAR model consistently
produces the lowest bias across all coefficients and data-generating
scenarios. This confirms its ability to recover true parameter values
under spatial dependency, even when data is aggregated. The SAR model
displays higher variability and often substantial bias, especially in
CAR-generated datasets, suggesting model misspecification when the true
structure is localised (as in CAR). GLM, which ignores spatial effects,
performs reasonably in some settings but often underperforms compared to
CAR.

In the point-level setting (Figure~\ref{fig-biasplot2}), the mlvCAR
model shows superior performance, achieving minimal bias across all
coefficients and data-generating mechanisms. This reflects its strength
in modeling spatial dependencies while accounting for multilevel data
structure (i.e., multiple observations within areas). GLMM performs
moderately well but tends to show slightly higher bias than mlvCAR. In
contrast, GWR, despite being designed for point-level spatial variation,
yields higher bias, especially under SAR+CAR structures. This suggests
GWR's limitation when spatial heterogeneity is not smooth or when model
assumptions are violated.

@@

Moreover, Table~\ref{tbl-rmse} provide insights into model accuracy at
different spatial levels. For area-level data, the models GLM, SAR, and
CAR are compared, with the CAR model consistently achieving the lowest
RMSE. This suggests that CAR is more effective at capturing spatial
dependencies at the area level. For point-level data, the models GLMM,
GWR, and mlvCAR are evaluated, with mlvCAR showing the lowest RMSE
across all datasets. This indicates that mlvCAR is particularly
well-suited for handling detailed spatial variations at the point level.
Overall, these results highlight the strengths of CAR and mlvCAR in
reducing error within their respective spatial contexts, underscoring
their potential utility in spatial modeling applications.

@@

In addition to the point-level dataset, which contains individual
property listings, the imputed dataset was also aggregated to the
sub-district level by calculating the mean values of key variables. This
aggregation produced an area-level dataset, allowing the analysis to be
conducted at both the individual and administrative levels. These two
levels of data granularity provide complementary perspectives for
evaluating the performance of spatial models in capturing local property
market dynamics.

@@

Standardization of the predictors ensures that the resulting
coefficients are on a comparable scale, facilitating meaningful
interpretation and comparison across different modeling approaches. The
results presented in Table~\ref{tbl-modelcomp} provide a comprehensive
evaluation of each model's performance in capturing both the structural
and spatial characteristics of the property market in Lombok.

We conducted model fitting on the area-level data, with detailed results
provided in the Appendix (Table~\ref{tbl-modelcomp1}). However, this
method does not provide optimal parameter estimates because data
aggregation can reduce variability and obscure spatial details essential
for capturing local effects. To address this, we used point-level data
to apply three different models: GLMM, GWR, and multilevel CAR.

@@

\begin{figure}

\begin{minipage}{0.50\linewidth}

\includegraphics{figures/phi_map.png}

\subcaption{\label{fig-pricenphi1}phi values}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\includegraphics{figures/logprice_map.png}

\subcaption{\label{fig-pricenphi2}Aggregate price data (log scale)}

\end{minipage}%

\caption{\label{fig-pricenphi}Heatmap of aggregated price data and phi
values in each Lombok sub-district}

Figure~\ref{fig-pricenphi1} and Figure~\ref{fig-pricenphi2} respectively
present spatial heatmaps illustrating the distribution of estimated
spatial random effects (\(\phi\)) from the CAR model and the average
log-transformed house prices per sub-district accross Lombok island.

Here, the \(\phi\) values represent the spatially structured residuals
not explained by the observed covariates in the CAR model. Specifically,
it captures latent spatial influences such as neighbourhood
desirability, access to infrastructure, or even local policy, that
affect house prices beyond what structural features like land and
building size can account for.

Because the model uses log-transformed prices, \(\phi\) has
multiplicative interpretation on the original price scale. For instance,
if \(\phi = 0.57\) for a given sub-district, then after accounting for
covariates, properties in that area are estimated to be approximately
\(exp(0.57) \approx 1.76\) or 76\% more expensive than average.
Similarly, a sub-district with \(\phi = -0.2\), this suggests that
properties in that area are approximately 18\% less expensive than
average, since \(exp(-0.2) \approx 0.82\).

@@

The findings from simulation consistently indicate that both the CAR and
mlvCAR model outperform the other models in capturing spatial
dependencies. This is evidenced by lower RMSE values and reduced bias
across various dataset scenarios, highlighting the CAR model's
robustness. In datasets with strong spatial structures, such as those
generated under \(CAR\) and \(SAR + CAR\) conditions, the CAR model
achieves substantially lower RMSE values compared to GLM and SAR,
suggesting its effectiveness in handling localised spatial variations.

@@

Spatial maps of the estimated random effects \(\phi\) from the CAR model
offer further insight into the spatial structure of housing prices
across Lombok Island. These effects represent residual spatial patterns
after accounting for structural covariates like land size, building
size, and number of rooms. Such interpretability supports the use of CAR
models not only for improved prediction but also for spatial
diagnostics. The visual comparison of \(\phi\) values and aggregated
log-transformed prices (as shown in Figure~\ref{fig-pricenphi})
highlights spatial patterns that may warrant further
investigation---such as clusters of residual over- or under-prediction
in West and East Lombok respectively. These patterns underscore the
added value of CAR-type models in revealing latent geographic effects.

These findings echo the limitations of GWR when precise coordinates are
lacking, affirming the relevance of area-level approaches. Moreover,
while we use a binary adjacency matrix for consistency, we recognise
that the choice of spatial weight \(\mathbf{W}\) may influences model
performance. Future studies should explore alternative specifications
such as row-standardized or distance-decay matrices to assess their
impact on robustness and interpretability.

Additionally, while the spatial dependence parameter \(\rho\) is
estimated as part of the model fitting, its variability across
simulations suggests that further sensitivity analyses---especially
under different fixed \(\rho\) values---could shed light on model
stability in different spatial contexts.

@@

In summary, this study highlights the importance of explicitly modeling
spatial dependence in house price analysis. CAR and mlvCAR models emerge
as practical and theoretically sound options, especially when
high-resolution spatial data are unavailable. Future research may extend
these models to finer spatial scales, incorporate additional spatial
diagnostics, or apply them in other domains where spatial heterogeneity
significantly influences outcomes.

@@

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}

\subsection*{Bias values for Area-level
Data}\label{bias-values-for-area-level-data}
\addcontentsline{toc}{subsection}{Bias values for Area-level Data}

@@

@@

\subsection*{Model Comparison for area-level
data}\label{model-comparison-for-area-level-data}
\addcontentsline{toc}{subsection}{Model Comparison for area-level data}

@@

\caption{\label{tbl-modelcomp1}This table compares the performance of
GLM, SAR, and CAR models on area-level data. It reports parameter
estimates (with 95\% confidence intervals), spatial parameters
(\(\rho\)), variances (\(\nu^2\), \(\tau^2\)), and model fit criteria
(AIC, DIC, WAIC, LMPL, log-likelihood). The results highlight each
model's capability to capture spatial dependencies and variability}

@@

\subsection*{Example of beta-values in GWR
model}\label{example-of-beta-values-in-gwr-model}
\addcontentsline{toc}{subsection}{Example of beta-values in GWR model}
